---
title: "A compilation of replications of the Diener et al.2010 paper"
author: "Theodore Rogers, Olivia Costa, Quincey Feragen, Penelope Corbett"
date: "2026-01-27"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
This is the README script to clean the replications from the Diener et al. (2010) project. R version 4.5.2 (2025-10-31) is used for this script. 

Packages needed:
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(tidyverse)
library(ltm)
library(gridExtra)
library(rmarkdown)
```

Reading excel spreadsheet
```{r setup, include=FALSE}
big_data <- read.csv("Diener_et_al_replication_data.csv")
#View(big_data)
```
creating data frame, the original data frame (big_data) is left untouched
```{r}
dat <- big_data
#view(dat)
dat = cbind.data.frame(big_data)
```

Changing values to their categorical variables
```{r}
#ladder
ladder <- as.factor(dat$ladder)
dat$ladder <- ladder

convert_yes_no <- function(column) {
  column <- as.character(column)
  column[column == "1"] <- "yes"
  column[column == "2"] <- "no"
  return(column)
}

dat[3:8] <- lapply(dat[3:8], convert_yes_no)
dat[11:16] <- lapply(dat[11:16], convert_yes_no)
dat[19:23] <- lapply(dat[19:23], convert_yes_no)

# gender
dat$gender[dat$gender == "1"] <- "woman"
dat$gender[dat$gender == "2"] <- "man"
dat$gender[dat$gender == "3"] <- "nonbinary"
dat$gender[dat$gender == "4"] <- NA
# maritalStatus
dat$maritalStatus[dat$maritalStatus == "1"] <- "married"
dat$maritalStatus[dat$maritalStatus == "2"] <- "domestic partnership"
dat$maritalStatus[dat$maritalStatus == "3"] <- "single"
dat$maritalStatus[dat$maritalStatus == "4"] <- "divorced or separated"
dat$maritalStatus[dat$maritalStatus == "5"] <- "widowed"
dat$maritalStatus[dat$maritalStatus == "6"] <- NA

# residentialArea
dat$residentialArea[dat$residentialArea == "1"] <- "rural"
dat$residentialArea[dat$residentialArea == "2"] <- "village"
dat$residentialArea[dat$residentialArea == "3"] <- "urban_city"
dat$residentialArea[dat$residentialArea == "4"] <- "suburban"

# education
dat$education[dat$education == "1"] <- "primary or less"
dat$education[dat$education == "2"] <- "secondary"
dat$education[dat$education == "3"] <- "some teriary"
dat$education[dat$education == "4"] <- "full teritary"
dat$education[dat$education == "5"] <- "graduate degree"
dat$education[dat$education == "6"] <- "other"

#Cleaning Income Data
dat$income <- as.character(dat$income)
dat$income[dat$income %in% c("NA", ".")] <- NA
dat$income <- as.numeric(dat$income)
```

**subsets for dataSets split by country:**
For this section, I modified the values of the different incomes by converting their income to 2023 USD, after looking up the conversion rate on Google. 
Recommendation: You may want to consider removing any outlier's in the data. Additionally, several dataframes were created by continent in case someone would like to analyze the data between continents. 
```{r}
dat <- dat %>%
  mutate(
    raw_income = readr::parse_number(as.character(income)),
    income = case_when(
      dataSet == "HongKong" ~ raw_income * 0.13,
      dataSet == "vienna"   ~ raw_income * 1.09,
      dataSet %in% c("toronto", "202psy") ~ raw_income * 0.73,
      TRUE ~ raw_income
    )
  )

#Sanity check to make sure no rows were missed in the above conversion. All good!
dat %>%
  filter(
    !dataSet %in% c(
      "HongKong", "vienna", "toronto", "202psy",
      "2018spring", "psy250.", "hatton", "haneda", "2019replication"
    )
  ) %>%
  count(dataSet)

#Sanity check income ranges
dat %>%
  group_by(dataSet) %>%
  summarize(
    n = n(),
    na_income = sum(is.na(income)),
    min_income = min(income, na.rm = TRUE),
    median_income = median(income, na.rm = TRUE),
    max_income = max(income, na.rm = TRUE)
  )
```

Cleaning the variables in the data frame, and coding them in accordance with the replication survey values.
```{r}
positive_vars <- c("enjoyment", "smileLaugh", "respect", "learn", "opportunity", "livingStandard", "emergency", "time", "food", "shelter", "internetAccess", "computer", "television")

#Recode function for yes/nmo -> 1/0 (positive)
for (var in positive_vars){
  vals <- unique(dat[[var]])
  
  if (any(c("yes", "no") %in% vals)) {
    dat[[var]] <- as.numeric(dplyr::recode(tolower(as.character(dat[[var]])), "yes" = 1, "no"= 0, .default = NA_real_))
  }else{
    message(paste("Skipping", var, "no-not in yes/no format"))
  }
}

#Recode function for yes/nmo -> -1/0 (positive)
negative_vars <- c("sadness", "anger", "depression", "worry")

for (var in negative_vars){
  vals <- unique(dat[[var]])
  
  if (any(c("yes", "no") %in% vals)) {
    dat[[var]] <- as.numeric(dplyr::recode(tolower(as.character(dat[[var]])), "yes" = -1, "no"= 0, .default = NA_real_))
  }else{
    message(paste("Skipping", var, "no-not in yes/no format"))
  }
}
```

Cleaning the variables in the data frame, and coding them in accordance with the replication survey values.
```{r}
#Average smileLaugh and enjoyment into Positive feelings column
dat$PositiveFeelings <- rowMeans(dat[, c("smileLaugh", "enjoyment")], na.rm = TRUE)

#Average worry, sadness, depression, and anger into Negative feelings column
dat$NegativeFeelings <- rowMeans(dat[, c("worry", "sadness", "depression", "anger")], na.rm = TRUE)

#Create another column with logIncome
dat$logIncome <- ifelse(!is.na(dat$income) & dat$income > 0,
                        log(dat$income),
                        NA_real_)

#Create another column of luxury conveniences
dat$LuxuryConveniences <- rowMeans(dat[, c("television", "internetAccess", "computer")], na.rm = TRUE)

#Create another column of basic needs
dat$BasicNeeds <- rowMeans(dat[, c("food", "shelter")], na.rm = TRUE)

#Create another column of psych needs
dat$PsychNeeds <- rowMeans(dat[, c("respect", "learn", "opportunity", "time", "emergency")], na.rm = TRUE)
```

Renaming study site variables for cleanliness
```{r}
dat <- dat %>%
  dplyr::mutate(
    study_site = dplyr::case_when(
      dataSet == "toronto" ~ "Study_Site_1",
      dataSet == "vienna" ~ "Study_Site_2",
      dataSet == "2018spring" ~ "Study_Site_3",
      dataSet == "2019replication" ~ "Study_Site_4",
      dataSet == "psy250." ~ "Study_Site_5",
      dataSet == "HongKong" ~ "Study_Site_6",
      dataSet == "hatton" ~ "Study_Site_7",
      dataSet == "202psy" ~ "Study_Site_8",
      dataSet == "haneda" ~ "Study_Site_9",
      TRUE ~ NA_character_
    )
  )
class(dat$study_site)
```

Demographics per study site made into tables
```{r}
#Sanity check on age 
dat <- dat %>%
  mutate(
    age = readr::parse_number(as.character(age))
  )

#Numeric demographics table (age & income)
numeric_demographics_table <- dat %>%
  dplyr::group_by(study_site) %>%
  dplyr::summarize(
    n = n(),

    age_n = sum(!is.na(age)),
    age_mean   = if (age_n > 0) mean(age, na.rm = TRUE) else NA_real_,
    age_median = if (age_n > 0) median(age, na.rm = TRUE) else NA_real_,
    age_min    = if (age_n > 0) min(age, na.rm = TRUE) else NA_real_,
    age_max    = if (age_n > 0) max(age, na.rm = TRUE) else NA_real_,

    income_n = sum(!is.na(income)),
    income_mean   = if (income_n > 0) mean(income, na.rm = TRUE) else NA_real_,
    income_median = if (income_n > 0) median(income, na.rm = TRUE) else NA_real_,
    income_min    = if (income_n > 0) min(income, na.rm = TRUE) else NA_real_,
    income_max    = if (income_n > 0) max(income, na.rm = TRUE) else NA_real_,

    .groups = "drop"
  )

#Categorical demographics table (counts & proportions)
categorical_demographics <- dat %>%
  dplyr::select(study_site, gender, maritalStatus, residentialArea, education) %>%
  pivot_longer(
    cols = c(gender, maritalStatus, residentialArea, education),
    names_to = "variable",
    values_to = "value"
  ) %>%
  dplyr::mutate(
    value = as.character(value),
    value = if_else(is.na(value) | trimws(value) == "", "Missing", value)
  ) %>%
  count(study_site, variable, value, name = "n") %>%
  group_by(study_site, variable) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  arrange(study_site, variable, desc(n))

#Which study sites are missing whole fields?
missingness <- dat %>%
  group_by(study_site) %>%
  summarize(
    age_missing = mean(is.na(age)),
    gender_missing = mean(is.na(gender)),
    marital_missing = mean(is.na(maritalStatus)),
    resarea_missing = mean(is.na(residentialArea)),
    education_missing = mean(is.na(education)),
    income_missing = mean(is.na(income)),
    .groups = "drop"
  ) %>%
  arrange(desc(age_missing), desc(income_missing))

#Viewing results
numeric_demographics_table
categorical_demographics
missingness
```

**Prepping for analyses:**
There will be two data frames, one with all participants (N = 1171), and the other that has all the NAs omitted across the dataframe. This was done, as a part of a previous project. That used a smaller dataset. 
```{r}
dat <- dat %>%
  dplyr::mutate(
    ladder = readr::parse_number(na_if(as.character(ladder), "NaN")),
    PositiveFeelings = readr::parse_number(na_if(as.character(PositiveFeelings), "NaN")),
    NegativeFeelings = readr::parse_number(na_if(as.character(NegativeFeelings), "NaN"))
  )

#Creating the omitted NAs data frame: this was used in the poster and other subsequent analyses, however, it is now its own data frame and you will be able to use Dat for the analyses below. 
dat_na_omit <- dat[complete.cases(dat), ] #N = 342 (whereas Dat has N = 1171)
```

**Sample Analyses**
For these analyses, I have dropped NAs in the columns we care about for these analyses and not overall throughout the whole data frame. To keep the sample consistent model to model.

Creating the models for the predictor variables (basic needs, psych needs, luxury conveniences, standard of living, and log income)
```{r}
f_basic  <- ladder ~ BasicNeeds
f_psych  <- ladder ~ BasicNeeds + PsychNeeds
f_lux    <- ladder ~ BasicNeeds + PsychNeeds + LuxuryConveniences + livingStandard
f_income <- ladder ~ BasicNeeds + PsychNeeds + LuxuryConveniences + livingStandard + logIncome

#Creating NA-datasets for each model
dat_basic  <- dat %>% drop_na(all.vars(f_basic))
dat_psych  <- dat %>% drop_na(all.vars(f_psych))
dat_lux    <- dat %>% drop_na(all.vars(f_lux))
dat_income <- dat %>% drop_na(all.vars(f_income))

#Make sure the N is locked in before running anything
data.frame(
  mode_spec = c("BasicNeeds", "PsychNeeds", "Luxury", "Income"),
  N = c(nrow(dat_basic), nrow(dat_psych), nrow(dat_lux), nrow(dat_income))
)

#Fitting models to the clean datasets
wellbeingBasicNeedsModel <- lm(f_basic,  data = dat_basic)
wellbeingPsychNeedsModel <- lm(f_psych,  data = dat_psych)
wellbeingLuxuryModel     <- lm(f_lux,    data = dat_lux)
wellbeingIncomeModel     <- lm(f_income, data = dat_income)

#Summaries
summary(wellbeingBasicNeedsModel)
summary(wellbeingPsychNeedsModel)
summary(wellbeingLuxuryModel)
summary(wellbeingIncomeModel)
```

Positivity Models
```{r}
#Defining the formulas for the positivity models
f_pos_basic  <- PositiveFeelings ~ BasicNeeds
f_pos_psych  <- PositiveFeelings ~ BasicNeeds + PsychNeeds
f_pos_lux    <- PositiveFeelings ~ BasicNeeds + PsychNeeds + LuxuryConveniences + livingStandard
f_pos_income <- PositiveFeelings ~ BasicNeeds + PsychNeeds + LuxuryConveniences + livingStandard + logIncome

#Create NA-clean datasets per model
dat_pos_basic  <- dat %>% drop_na(all.vars(f_pos_basic))
dat_pos_psych  <- dat %>% drop_na(all.vars(f_pos_psych))
dat_pos_lux    <- dat %>% drop_na(all.vars(f_pos_lux))
dat_pos_income <- dat %>% drop_na(all.vars(f_pos_income))

#Fitting the models on the clean datasets
positiveBasicNeedsModel <- lm(f_pos_basic,  data = dat_pos_basic)
positivePsychNeedsModel <- lm(f_pos_psych,  data = dat_pos_psych)
positiveLuxuryModel     <- lm(f_pos_lux,    data = dat_pos_lux)
positiveIncomeModel     <- lm(f_pos_income, data = dat_pos_income)

#Clearing N Table
data.frame(
  model_spec = c("BasicNeeds",
                 "BasicNeeds + PsychNeeds",
                 "BasicNeeds + PsychNeeds + Luxury + LivingStandard",
                 "Full + logIncome"),
  N = c(nrow(dat_pos_basic),
        nrow(dat_pos_psych),
        nrow(dat_pos_lux),
        nrow(dat_pos_income))
)

#Summaries
summary(positiveBasicNeedsModel)
summary(positivePsychNeedsModel)
summary(positiveLuxuryModel)
summary(positiveIncomeModel)
```

Negativity Models
```{r} 
#Defining the formulas for the negativity models
f_neg_basic  <- NegativeFeelings ~ BasicNeeds
f_neg_psych  <- NegativeFeelings ~ BasicNeeds + PsychNeeds
f_neg_lux    <- NegativeFeelings ~ BasicNeeds + PsychNeeds + LuxuryConveniences + livingStandard
f_neg_income <- NegativeFeelings ~ BasicNeeds + PsychNeeds + LuxuryConveniences + livingStandard + logIncome

#Create NA-clean datasets per model
dat_neg_basic  <- dat %>% drop_na(all.vars(f_neg_basic))
dat_neg_psych  <- dat %>% drop_na(all.vars(f_neg_psych))
dat_neg_lux    <- dat %>% drop_na(all.vars(f_neg_lux))
dat_neg_income <- dat %>% drop_na(all.vars(f_neg_income))

#Fitting the models on the clean datasets
negativeBasicNeedsModel <- lm(f_neg_basic,  data = dat_neg_basic)
negativePsychNeedsModel <- lm(f_neg_psych,  data = dat_neg_psych)
negativeLuxuryModel     <- lm(f_neg_lux,    data = dat_neg_lux)
negativeIncomeModel     <- lm(f_neg_income, data = dat_neg_income)

#Clearing N Table
data.frame(
  model_spec = c("BasicNeeds",
                 "BasicNeeds + PsychNeeds",
                 "BasicNeeds + PsychNeeds + Luxury + LivingStandard",
                 "Full + logIncome"),
  N = c(nrow(dat_neg_basic), nrow(dat_neg_psych), nrow(dat_neg_lux), nrow(dat_neg_income))
)

#Summaries
summary(negativeBasicNeedsModel)
summary(negativePsychNeedsModel)
summary(negativeLuxuryModel)
summary(negativeIncomeModel)
```


Income Demographics
```{r}
#Total Income
total_income <- dat %>%
  summarise(
    n_total = n(),
    n_income = sum (!is.na(income)),
    percent_income = mean(!is.na(income))*100,
    mean_income = mean(income, na.rm = TRUE),
    median_income = median(income, na.rm = TRUE),
    sd_income = sd(income, na.rm = TRUE),
    min_income = min(income, na.rm = TRUE),
    max_income = max(income, na.rm = TRUE),
    q1 = quantile(income, .25, na.rm = TRUE),
    q3 = quantile(income, .75, na.rm = TRUE)
  )

# Total Income by study site
income_by_site <- dat %>%
  group_by(study_site)%>%
  summarise(
    n_total = n(),
    n_income = sum (!is.na(income)),
    percent_income = mean(!is.na(income))*100,
    mean_income = mean(income, na.rm = TRUE),
    median_income = median(income, na.rm = TRUE),
    sd_income = sd(income, na.rm = TRUE),
    min_income = min(income, na.rm = TRUE),
    max_income = max(income, na.rm = TRUE),
    q1 = quantile(income, .25, na.rm = TRUE),
    q3 = quantile(income, .75, na.rm = TRUE)
  )
```

Graphs
```{r}
#graph of positive feelings ~ Basic Needs
graphDat <- dat

ggplot(graphDat, aes(PositiveFeelings))+
  geom_histogram(aes(BasicNeeds, fill = "Basic Needs"), binwidth = 0.5, color = "black", alpha = 0.7)+
  geom_histogram(aes(PsychNeeds, fill = "Psychological Needs"), binwidth = 0.5, color = "black", alpha = 0.7)+
  theme_classic()+
  labs(title = "Relationship between Postive Feelings, Psychological Needs, and Basic Needs", 
       y = "Participants", 
       x = "Positive Feelings")+
  scale_fill_manual(values = c("Basic Needs" = "darkgoldenrod3", "Psychological Needs" = "firebrick"), 
                    name = "Variables")

##graphing this another way
ggplot(graphDat, aes(PositiveFeelings, PsychNeeds))+
  geom_jitter(color = "firebrick4")+
  geom_smooth(method = "lm", se = FALSE, color = "black")+
  theme_classic()+
  labs(title = "Relationship between Positive Feelings and Psychological Needs", 
       x = "Positive Feelings", 
       y = "Psychological Needs")+ theme(text = element_text(size = 14))

#graph of Income and Positive Feelings
ggplot(graphDat, aes(PositiveFeelings, logIncome))+
  geom_jitter()+
  geom_smooth(method = "lm", se = FALSE, color = "maroon")+
  theme_classic()+
  labs(title = "Relationship between Log Income and Positive Feelings", 
       x = "Positive Feelings", 
       y = "Log Income")+ theme(text = element_text(size = 14))

#graph of Negative Feelings and Income
ggplot(graphDat, aes(NegativeFeelings, logIncome))+
  geom_jitter()+
  geom_smooth(method = "lm", se = FALSE, color = "darkgoldenrod3")+
  theme_classic()+
  labs(title = "Relationship between Log Income and Negative Feelings", 
       x = "Negative Feelings", 
       y = "Log Income")+ theme(text = element_text(size = 14))

#graph of Positive feelings and luxury items
ggplot(graphDat, aes(PositiveFeelings, LuxuryConveniences))+
  geom_jitter(color = "firebrick4")+
  geom_smooth(method = "lm", se = FALSE, color = "black")+
  theme_classic()+
  labs(title = "Relationship between Positive Feelings and Luxury Conveniences", 
       x = "Positive Feelings", 
       y = "Luxury Conveniences")+ theme(text = element_text(size = 14))

```

